---
title: "(Pseudo)Legal Heatmaps"
date: '2024-02-14'
output: html_document
toc: true
toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Legal and Pseudolegal Documents Heatmaps

## Setup 
This RMD has all the code for creating the heatmaps. 
We start by setting up the workplace.

```{r setting_up, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
# libraries
library(tidyverse) # for the general graphs
library(dplyr) # for handling the data
library(MASS) # for density estimation
library(colorRamp2) # to generate color palettes
library(spatstat.geom) # for geometrical maths

# setup math display options
options(scipen=9)

# read in data
LLD <- read.csv(
  file = '/Users/dana/Documents/R/heatmaps_lal/LLD_All_Features_clean.csv', 
  header = TRUE) 
SCD <- read.csv(
  file = '/Users/dana/Documents/R/heatmaps_lal/SCD_All_Features_clean.csv', 
  header = TRUE) 

# slashes create problems when automatically naming files,
# so we remove / in the feature names
LLD$feature <- gsub("\\/", 
                    "", 
                    LLD$feature)
SCD$feature <- gsub("\\/", 
                    "", 
                    SCD$feature)

# this gets unique features, counts occurrences, drops any under 10
LLD_list_df <- LLD %>% 
  group_by(feature) %>% 
  summarize(count=n()) %>% 
  filter(count >= 10)
SCD_list_df <- SCD %>% 
  group_by(feature) %>% 
  summarize(count=n()) %>% 
  filter(count >= 10)

# for maths, different list to loop is needed
LLD_list_names <- as.data.frame(unique(LLD$feature))
SCD_list_names <- as.data.frame(unique(SCD$feature))
colnames(LLD_list_names) <- "feature"
colnames(SCD_list_names) <- "feature"

# find intersection of the feature lists
both_list <- intersect(LLD_list_names, 
                   SCD_list_names)

# for the maths bit, we need functions to convert between measurements
# acos works with radians, not with degrees
# https://stackoverflow.com/questions/32370485/convert-radians-to-degree-degree-to-radians
deg2rad = function(deg) {
  return((pi * deg) / 180)
}

rad2deg = function(rad) {
  return((180 * rad) / pi)
}
```

## Plot Creation

Then we can automatically create all the relevant plots and output them in a folder for all features. We split this up in two chunks, one for the legal and one for the pseudolegal documents. 
This code loops through the list of features that were coded originally. Then it calculates a probability density estimation, in this case a kernel density estimation of the spatial dimensions of the feature on the page.
Two maps or graphs per feature are created. The first is a histogram of the underlying (original) data mapped on the page, so we can understand where each feature predominantly occurs. The second graph is a visualisation of the kernel density estimation, again mapped on the page. 

```{r plots_lld}
# loop for LLD
# for each feature in our list of features
for (feature_item in LLD_list_df$feature) {
  # get feature
  LLD_feature <- LLD %>% 
    dplyr::filter(feature == feature_item)
  
  # get the coordinates of the feature
  LLD_feature_coord <- LLD_feature %>% 
    dplyr::select(x, y) %>% 
    as.data.frame()
  
  # histogram output in jpeg format
  jpeg(
    paste0("/Users/dana/Documents/R/heatmaps_lal/first_outputs/LLD_histogram_", 
           feature_item,".jpeg"), 
    units = "in", 
    width = 6, 
    height = 7.5, 
    res = 300)
  print(
    # this is the actual graph
    ggplot(LLD_feature_coord, 
           aes(x = x, y = y)) +
    geom_hex(bins = 30, 
             show.legend = FALSE) +
    scale_fill_viridis(option = "magma", 
                       direction = -1) +
    theme_bw() +
    xlab(" ") +
    ylab(" ") +
    scale_x_continuous(expand=c(0, 0), 
                       limits=c(0, 1700)) +
    scale_y_continuous(expand=c(0, 0), 
                       limits=c(2210, 0), 
                       trans = "reverse") +
    ggtitle(paste("LLD: ", 
                  feature_item),
            subtitle = "Histogram"))
  dev.off()

  # define colours for heatmap
  mycolors <- colorRampPalette(c("white", "#cc0000"))(250)
  
  # density plot output in jpeg format
  jpeg(
    paste0("/Users/dana/Documents/R/heatmaps_lal/first_outputs/LLD_density_", 
           feature_item,".jpeg"), 
    units = "in", 
    width = 6, 
    height = 7.5, 
    res = 300)
  print(
    # this is the actual graph
    ggplot(LLD_feature_coord, 
           aes(x, y)) +
      stat_density_2d_filled(geom = "density_2d_filled",
                             contour_var = "ndensity",
                             n = 50,
                             bins = 250,
                             show.legend = FALSE) +
      scale_fill_manual(values = mycolors,
                        na.value = "transparent") +
  theme_bw() +
  xlab(" ") +
  ylab(" ") +
  scale_x_continuous(expand=c(0, 0), 
                     limits=c(0, 1700)) +
  scale_y_continuous(expand=c(0, 0), 
                     limits=c(2210, 0), 
                     trans = "reverse") +
    ggtitle(paste("LLD: ", 
                  feature_item),
            subtitle = "Density Plot"))
  dev.off()
}
```

Same as above, but for the other subcorpus.

```{r plots_scd}
# loop for SCD
# for each feature in our list of features
for (feature_item in SCD_list_df$feature) {
  # get feature
  SCD_feature <- SCD %>% 
    filter(feature == feature_item)
  
  # get the coordinates of the feature
  SCD_feature_coord <- SCD_feature %>% 
    dplyr::select(x, y) %>% 
    as.data.frame()

  # histogram output in jpeg format
  jpeg(
    paste0("/Users/dana/Documents/R/heatmaps_lal/first_outputs/SCD_histogram_", 
           feature_item,".jpeg"), 
    units = "in", 
    width = 6, 
    height = 7.5, 
    res = 300)
  print(
    # this is the actual graph
    ggplot(SCD_feature_coord, 
           aes(x = x, y = y)) +
    geom_hex(bins = 30, 
             show.legend = FALSE) +
    scale_fill_viridis(option = "magma", 
                       direction = -1) +
    theme_bw() +
    xlab(" ") +
    ylab(" ") +
    scale_x_continuous(expand=c(0, 0), 
                       limits=c(0, 1700)) +
    scale_y_continuous(expand=c(0, 0), 
                       limits=c(2210, 0), 
                       trans = "reverse") +
    ggtitle(paste("SCD: ", 
                  feature_item),
            subtitle = "Histogram"))
  dev.off()
  
  # define colours for heatmap
  mycolors <- colorRampPalette(c("white", "#cc0000"))(250)

  # density plot output in jpeg format
  jpeg(
    paste0("/Users/dana/Documents/R/heatmaps_lal/first_outputs/SCD_density_", 
           feature_item,".jpeg"), 
    units = "in", 
    width = 6, 
    height = 7.5, 
    res = 300)
  print(
    # this is the actual graph
    ggplot(SCD_feature_coord, 
           aes(x, y)) +
      stat_density_2d_filled(geom = "density_2d_filled",
                             contour_var = "ndensity",
                             n = 50,
                             bins = 250,
                             show.legend = FALSE) +
      scale_fill_manual(values = mycolors,
                        na.value = "transparent") +
    theme_bw() +
    xlab(" ") +
    ylab(" ") +
    scale_x_continuous(expand=c(0, 0), 
                       limits=c(0, 1700)) +
    scale_y_continuous(expand=c(0, 0), 
                       limits=c(2210, 0), 
                       trans = "reverse") +
    ggtitle(paste("SCD: ", 
                  feature_item),
            subtitle = "Density Plot"))
  dev.off()
}
```

## Single Plots

We can also just create single plots and not automate the creation by looping through the features. This also allows for tweaking the visualisation if necessary. First is the Histogram. 

```{r single_histogram}
# filter the list for the item of interest
LLD_feature <- LLD %>% 
  filter(feature == "footer")
# get coordinates for item of interest
LLD_feature_coord <- LLD_feature %>% 
  dplyr::select(x, y) %>% 
  as.data.frame()

# filter the list for the item of interest
SCD_feature <- SCD %>% 
  filter(feature == "footer")
# get coordinates for item of interest
SCD_feature_coord <- SCD_feature %>% 
  dplyr::select(x, y) %>% 
  as.data.frame()

# plotting of the feature
## change the first argument in ggplot() to 
## reflect the respective subcorpus
ggplot(SCD_feature_coord, 
       aes(x = x, y = y)) +
  geom_hex(bins = 30, 
           show.legend = FALSE) +
  scale_fill_viridis(option = "magma", 
                     direction = -1) +
  theme_bw() +
  xlab(" ") + 
  ylab(" ") +
  scale_x_continuous(expand=c(0, 0), 
                     limits=c(0, 1700)) +
  scale_y_continuous(expand=c(0, 0), 
                     limits=c(2210, 0), 
                     trans = "reverse") +
  ggtitle("SCD: Footer",
          subtitle = "Histogram") 
```

Second is the density plot outside of the loop.

```{r single_density_plot}
# define colours for heatmap
mycolors <- colorRampPalette(c("white", "#cc0000"))(250)

## change the first argument in ggplot() to 
## reflect the respective subcorpus
ggplot(LLD_feature_coord, 
           aes(x, y)) +
      stat_density_2d_filled(geom = "density_2d_filled",
                             contour_var = "ndensity",
                             n = 50,
                             bins = 250,
                             show.legend = FALSE) +
      scale_fill_manual(values = mycolors,
                        na.value = "transparent") +
    theme_bw() +
    xlab(" ") +
    ylab(" ") +
    scale_x_continuous(expand=c(0, 0), 
                       limits=c(0, 1700)) +
    scale_y_continuous(expand=c(0, 0), 
                       limits=c(2210, 0), 
                       trans = "reverse") +
    ggtitle("SCD: Footer",
            subtitle = "Density Plot")
```

## Comparison

If we want to mathematically compare which feature differs the most between SCD and LLD, we can calculate the angle between the features. 
In order to do this, a kernel density estimation is run for a feature for both subcorpora. Then the mean density point is calculated and added to a matrix. With this, we have three points, (0,0) and the two density estimation means. They form a triangle, with a different size angle at the zero point of the graph. We calculate this angle for each feature and append it to a matrix, which we then sort.

```{r angle_calc}
# nrow needs to be the number of features
results_matrix <- as.data.frame(matrix(NA, 
                                       ncol = 1, 
                                       nrow = 35))
colnames(results_matrix) <- "distance"
rownames(results_matrix) <- both_list$feature

# loop over all features
for (feature_item in both_list$feature) {
  # create temporary feature variable
  temp_feature <- feature_item
  
  # get feature and coordinates from dataframes
  ## same as above
  LLD_feature <- LLD %>% 
    dplyr::filter(feature == feature_item)
  LLD_feature_coord <- LLD_feature %>% 
    dplyr::select(x, y) %>% 
    as.data.frame()
  SCD_feature <- SCD %>% 
    filter(feature == feature_item)
  SCD_feature_coord <- SCD_feature %>% 
    dplyr::select(x, y) %>% 
    as.data.frame()
  
  # density calculation and transformation into df
  ## above this happens within ggplot2
  density_LLD <- kde2d(LLD_feature_coord$x, 
                       LLD_feature_coord$y,
                       h = 50)
  density_LLD_df <- as.data.frame(density_LLD)
  density_SCD <- kde2d(SCD_feature_coord$x, 
                       SCD_feature_coord$y,
                       h = 50)
  density_SCD_df <- as.data.frame(density_SCD)
  
  # set up matrix for distance calculation
  points <- matrix(0, 
                   ncol = 3, 
                   nrow = 2)
  rownames(points) <- c("x", 
                        "y")
  colnames(points) <- c("start", 
                        "lld", 
                        "scd")
  
  # assign the mean of the coordinate points from density output
  ## since we require two pairs of coordinates, we need to average over 
  ## 4 coordinate points in total
  points[1,2] <- mean(density_LLD_df$x, 
                      trim = 0, 
                      na.rm = FALSE)
  points[2,2] <- mean(density_LLD_df$y, 
                      trim = 0, 
                      na.rm = FALSE)
  points[1,3] <- mean(density_SCD_df$x, 
                      trim = 0, 
                      na.rm = FALSE)
  points[2,3] <- mean(density_SCD_df$y, 
                      trim = 0, 
                      na.rm = FALSE)
  
  # calculate distances between the points
  ## 3 calculations, because we have three points / three lines
  distance_start_lld = crossdist(points[1,1], 
                                 points[2,1], 
                                 points[1,2], 
                                 points[2,2]) 
  distance_start_scd = crossdist(points[1,1], 
                                 points[2,1], 
                                 points[1,3], 
                                 points[2,3]) 
  distance_lld_scd = crossdist(points[1,2], 
                               points[2,2], 
                               points[1,3], 
                               points[2,3]) 

  # this calculates the angle between SCD and LLD
  # https://math.stackexchange.com/questions/292935/how-do-i-find-the-angles-of-a-triangle-if-i-only-have-the-lengths-of-the-sides
  number_calc <- ((distance_start_scd^2) + 
                    (distance_start_lld^2) - 
                    (distance_lld_scd^2)) / 
    (2*distance_start_scd*distance_start_lld)
  angle_radian <- acos(number_calc)
  
  # convert to degrees
  angle_degree <- rad2deg(angle_radian)
  
  # assign output to matrix
  results_matrix[temp_feature, 1] <- angle_degree
  
}

# sort the results  
results_zerobase <- results_matrix %>% 
  arrange(desc(distance))



## second loop with different reference point
# loop over all features
for (feature_item in both_list$feature) {
  # create temporary feature variable
  temp_feature <- feature_item
  
  # get feature and coordinates from dataframes
  ## same as above
  LLD_feature <- LLD %>% 
    dplyr::filter(feature == feature_item)
  LLD_feature_coord <- LLD_feature %>% 
    dplyr::select(x, y) %>% 
    as.data.frame()
  SCD_feature <- SCD %>% 
    filter(feature == feature_item)
  SCD_feature_coord <- SCD_feature %>% 
    dplyr::select(x, y) %>% 
    as.data.frame()
  
  # density calculation and transformation into df
  ## above this happens within ggplot2
  density_LLD <- kde2d(LLD_feature_coord$x, 
                       LLD_feature_coord$y,
                       h = 50)
  density_LLD_df <- as.data.frame(density_LLD)
  density_SCD <- kde2d(SCD_feature_coord$x, 
                       SCD_feature_coord$y,
                       h = 50)
  density_SCD_df <- as.data.frame(density_SCD)
  
  # set up matrix for distance calculation
  points <- matrix(0, 
                   ncol = 3, 
                   nrow = 2)
  rownames(points) <- c("x", 
                        "y")
  colnames(points) <- c("start", 
                        "lld", 
                        "scd")
  
  # assign the mean of the coordinate points from density output
  ## since we require two pairs of coordinates, we need to average over 
  ## 4 coordinate points in total
  points[1,2] <- mean(density_LLD_df$x, 
                      trim = 0, 
                      na.rm = FALSE)
  points[2,2] <- mean(density_LLD_df$y, 
                      trim = 0, 
                      na.rm = FALSE)
  points[1,3] <- mean(density_SCD_df$x, 
                      trim = 0, 
                      na.rm = FALSE)
  points[2,3] <- mean(density_SCD_df$y, 
                      trim = 0, 
                      na.rm = FALSE)
  # this assigns the different reference point
  points[1,1] <- 0
  points[2,1] <- 2200
  
  # calculate distances between the points
  ## 3 calculations, because we have three points / three lines
  distance_start_lld = crossdist(points[1,1], 
                                 points[2,1], 
                                 points[1,2], 
                                 points[2,2]) 
  distance_start_scd = crossdist(points[1,1], 
                                 points[2,1], 
                                 points[1,3], 
                                 points[2,3]) 
  distance_lld_scd = crossdist(points[1,2], 
                               points[2,2], 
                               points[1,3], 
                               points[2,3]) 
  
  # this calculates the angle between SCD and LLD
  # https://math.stackexchange.com/questions/292935/how-do-i-find-the-angles-of-a-triangle-if-i-only-have-the-lengths-of-the-sides
  number_calc <- ((distance_start_scd^2) + 
                    (distance_start_lld^2) - 
                    (distance_lld_scd^2)) / 
    (2*distance_start_scd*distance_start_lld)
  angle_radian <- acos(number_calc)
  
  # convert to degrees
  angle_degree <- rad2deg(angle_radian)
  
  # assign output to matrix
  results_matrix[temp_feature, 1] <- angle_degree
  
}

# sort the results  
results_2kbase <- results_matrix %>% 
  arrange(desc(distance))

results_average <- merge(results_zero, 
                         results_2200, 
                         by = "feature")
colnames(results_average) <- c("feature", 
                               "angle_1", 
                               "angle_2")
results_average <- results_average %>% 
  rowwise() %>% 
  mutate(mean = mean(c(angle_1, angle_2)))

results_average <- results_average %>% 
  arrange(desc(mean))
```
